---
title: "Examining CV Data"
output: 
  html_document: 
    toc: yes
---

```{r setup, echo=FALSE, warning = FALSE, message = FALSE}
# source the file that contains the functions
source("cvFunctions.R")
# use cvSim to simulate a cyclic voltammogram accepting default options
testCV = cvSim()
# add the DT and magrittr packages for use in this file
library(DT)
library(magrittr)
```

When we use the function `cvSim` to create an object it returns to the object a list that contains the following vectors and matrices

* `time`: a vector that contains the discrete times used to define the diffusion grid
* `distance`: a vector that contains the discrete distances used to define the diffusion grid
* `potential`: a vector that contains the discrete applied potentials corresponding to the discrete times used in the simulation
* `current`: a vector that contains the calculated current at each discrete applied potential
* `oxdata`: a matrix that contains the concentrations of the oxidized species in the diffusion grid; each row in the matrix is a discrete time and each column in the matrix is a discrete distance 
* `reddata`: a matrix that contains red concentrations of the reduced species in diffusion grid; each row in the matrix is a discrete time and each column in the matrix is a discrete distance
* `formalE`: the redox couple's formal potential

We can access this data to examine the times, potentials, and currents generated by a simulation. All the functions described here are included in the file `cvFunctions.R`.

### Creating Tables to Examine the Data

A table is a convenient way to examine the relationship between several variables. To create a table, we first use the `tableCV` function

```
tableCV(file)
```

which takes as its only input the name of the object created by `cvSim`; thus

```{r create_table}
tableCV(testCV)
```

The resulting table displays the data across several pages, showing, by default, 10 rows per page. Clicking on the name of a column allows you to reorder the data by its rank.

### Finding Maximum Peak Currents and Peak Potentials

Although we can use the table created by the `displayTable` function to find the maximum peak currents and peak potentials, it is easier to use the `summaryCV` function 

```
summaryCV(file)
```

which takes as its only input the name of the object created by `cvSim`; thus

```{r summary}
summaryCV(testCV)
```

Note: Although imax corresponds to *i*~p,c~, imin does not correspond to *i*~p,a~ as it does not correct for the background current.