---
title: "Functions to Plot Results From cvSim"
output:
  html_document:
    toc: yes
---

```{r setup, echo=FALSE, warning = FALSE, message = FALSE}
# source the file that contains the functions
source("cvFunctions.R")
# use cvSim to simulate two cyclic voltammograms
testCV1 = cvSim()
testCV2 = cvSim(scan.rate = 0.16)
# add the animation library for use in this file
library(animation)
# add the plotly and ggplot2 libraries for use in this file
library(plotly)
library(ggplot2)
```

When we use the function `cvSim` to create an object it returns to the object a list that contains the following vectors and matricies

* `time`: a vector that contains the discrete times used to define the diffusion grid
* `distance`: a vector that contains the discrete distances used to define the diffusion grid
* `potential`: a vector that contains the discrete applied potentials corresponding to the discrete times used in the simulation
* `current`: a vector that contains the calulated current at each discrete applied potential
* `oxdata`: a matrix that contains the concentrations of the oxidized species in the diffusion grid; each row in the matrix is a discrete time and each column in the matrix is a discrete distance 
* `reddata`: a matrix that contains red concentrations of the reduced species in diffusion grid; each row in the matrix is a discrete time and each column in the matrix is a discrete distance
* `ormalE`: the redox couple's formal potential

We can access this data to display the results of a simulation in a variety of useful ways. All the functions described here are included in the document `CV Functions`.

### Plotting the Cyclic Voltammogram

To plot the cyclic voltammogram, we use the `plotCV` function 

```
plotCV(file, overlay = FALSE)
```

which takes the inputs shown in parentheses where `file` is the name of the object created by `cvSim` and `overlay` allows for plotting more than one CV on the same set of axes (note: the default option of FALSE draws the axes and the CV; setting `overlay` to TRUE plots the CV only). In addition to the cyclic voltammogram, the resulting output includes a vertical dashed line at the redox couple's formal potential and a horizontal dashed line at a current of zero. The following code shows how to overlay the results of two simulations: 

```{r plotCV}
# first, plot the results for testCV2
plotCV(testCV2)
# and then overlay the results for testCV1
plotCV(testCV1, overlay = TRUE)
```

We also can plot the cyclic voltammogram using the `plotly` library using the function

```
plotlyCV(file)
```

which takes as its only input the name of the object created by `cvSim`. The resulting ouput provides an interactive plot; move your cursor inside the plot to explore the possibilities.

```{r plotly}
plotlyCV(testCV1)
```

### Plotting the Applied Potential Waveform

To plot the applied potential waveform, we use the `plotPotential` function 

```
plotPotential(file)
```

which takes as its only input the name of the object created by `cvSim`. The resulting output is self-explantory.

```{r plotPotential}
plotPotential(testCV1)
```

### Ploting the Diffusion Profile at a Specific Time

A diffusion profile shows how the concentrations of the redox couple's oxidized and reduced forms change as a function of distance from the electrode surface at a particular time during cyclic voltammogram. To plot the diffusion profile, we use the `plotDiffusion` function 

```
plotDiffusion(file, index = 1, both_species = TRUE) 
```

which takes the inputs shown in parentheses where `file` is the name of the object created by `cvSim`, `index` is used to specify the time (for example, index = 1 is the first time unit), and `both_species` plots the diffusion profile for both the analyte's oxidized and its reduced forms (note: the default option of TRUE includes both; setting `both_speices` to FALSE plots the diffusion profile for the analyte's oxidized form only). The resulting output displays the concentration of the redox couple's oxidized form in blue and the concentration of its reduced form in red.

```{r plotDiffusion}
# let's examine the diffusion profile for the 100th point in simulation
plotDiffusion(testCV1, index = 100)
```

### Plotting a Grid of Diffusion Profiles

The function `plotGrid` produces a 3 $\times$ 3 grid of plots showing, at the grid's center, a cyclic voltammogram, and, arranged around the CV, eight diffusion profiles equally spaced in time. The function takes the form 

```
plotGrid(file)
```

which takes as its only input the name of the object created by `cvSim`. The following plot shows typical example.

```{r plotGrid, fig.width=7, fig.height=8}
plotGrid(testCV1)
```
